Vagueness, Noise, and Signaling

### 2015-08-11 paper/version_04/paper.tex
started on version 4

\newcommand{\rd}{\acro{rd}} % replicator dynamic
\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator

@@

@@

\begin{abstract}
  Signaling games are popular models for studying the evolution of meaning, but typical
  approaches do not incorporate vagueness as a feature of successful signaling.  Complementing
  recent like-minded models, we introduce the replicator diffusion dynamic that combines
  evolutionary pressure on optimal signal use with stochastic noise in the processing of
  similar stimuli. Applying this new dynamic to a generalization of Lewis' signaling games, we
  show that stochastic imprecision leads to vague, yet by-and-large efficient signal use, and,
  moreover, that it unifies evolutionary outcomes and helps avoid suboptimal categorization.
\end{abstract}

@@

A number of authors have recently tried to explain why vagueness evolved as something that is
itself useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
Others have argued that vagueness is a natural byproduct of limitations in information
processing or particular learning strategies
\citep[e.g.][]{FrankeJager2010:Vagueness-Signa,OConnor2013:The-Evolution-o}. This paper
contributes to the latter line of thought. Concretely, we introduce the replicator diffusion
dynamic that integrates confusability of similar stimuli. The dynamic proposed here generalizes
and complements previous like-minded accounts. We show that stochastic noise can not only lead
to vague, yet relatively efficient signal use, but can also unify evolutionary outcomes and
help avoid suboptimal categorization.

The next section introduces the background against which the work presented here can be
appreciated. Section~\ref{sec:repl-diff-dynam} introduces the replicator diffusion dynamic,
which is then explored in Section~\ref{sec:exploring-rdd}.  Section~\ref{sec:discussion}
reflects and compares our approach to others. Two appendices cover technical material that is
not strictly necessary to appreciate the main conceptual points of the paper.
Appendix~\ref{sec:diffusion-as-special} shows that our replicator diffusion dynamic is a
special case of the well-known replicator mutator dynamic. Appendix~\ref{sec:deriv-rdd-imit}
gives a derivation of our dynamic as the most-likely path of a population whose agents imitate
successful behavior without being aware of perceptual noise themselves.

@@

\section{Replicator diffusion dynamic}
\label{sec:repl-diff-dynam}

We introduce a population-level dynamic in which communicatively more
efficient strategies increase in proportion, as usual, but where the
realization of signaling strategies is noisy because of the agents'
inability to sharply distinguish similar states. We call our dynamic
\emph{replicator diffusion dynamic} (\rdd) because it is an extension
of the replicator dynamic (\rd) and a special case of the replicator
mutator dynamic (\rmd).

The formulation of the \rdd given and explored here is in terms of behavioral strategies, not
in terms of mixed strategies. Dynamics on behavioral strategies assume that agents can adapt
their behavior locally, so to speak, i.e., independently at each choice
point. Appendix~\ref{sec:repl-mutat-dynam} elaborates on the
difference. Appendix~\ref{sec:relating-dynamics} shows how to translate our initial formulation
of the \rdd into one using mixed strategies. Doing so identifies the \rdd as a special case of
the \rmd.

Our focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
numerical simulations. But it also points towards an interpretation of the process as cultural
evolution, rather than biological evolution. Signaling strategies are not inherited as a whole,
but adapted piece-meal, e.g., by conditional imitation of ``local'' behavior at a given choice
point. Appendix~\ref{sec:deriv-rdd-imit} gives one possible derivation of our dynamic as the
most likely path of development in a population of agents that imitate successful behavior
while being entirely unaware of perceptual imprecision themselves. Still, the possibility to
formulate the \rdd as a special case of the \rmd (Appendix~\ref{sec:diffusion-as-special}) also
suggests that an interpretation as a process of biological evolution may be feasible.

We begin by recapitulating the \rd in Section~\ref{sec:repl-dynam-behav}, and introduce
the \rdd in Section~\ref{sec:repl-diff-dynam-1}.
%Section~\ref{sec:repl-mutat-dynam} introduces the \rmd, and
%Section~\ref{sec:diffusion-as-special} shows in what sense the \rdd is
%a special case of the \rmd. Sections~\ref{sec:repl-mutat-dynam} and
%\ref{sec:diffusion-as-special} are conceptually important, but rather
%technical and not strictly necessary to appreciate the main point of
%the paper.

@@

Fix a signaling game with finite states $\States$, messages $\Messgs$
and acts $\Acts$. Let $\Pr(\cdot) \in \Delta{\States}$ be the prior
distribution over states and $\Util_{\sen,\rec} \mycolon \States
\times \Messgs \times \Acts \rightarrow \mathds{R}$ the sender's and
receiver's utility functions. A behavioral strategy is a function that
maps an agent's choice points to a probability distribution over
available choices. The sender's behavioral strategies are functions
$\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$ to a probability of each message $\messg \in \Messgs$ being sent; the receiver's are functions
$\Rstrat \in \Delta(\Acts)^\Messgs$, thus mapping each message $\messg \in \Messgs$ to a probability of each state $\state \in \States$ being interpreted by the receiver given $\messg$. The expected utility of choices
at each choice point is:
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
  \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
    \States} \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \cdot
  U_\rec(\state, \messg, \act) \,.
\end{align*}
This means that the expected utility for the sender of choosing message $\messg \in \Messgs$,
given tate $\state \in \States$ and a receiver strategy $\Rstrat$ is the sum, for each act
$\act \in \Acts$, of the probability that the receiver will choose act $\act$ given message
$\messg$ times the actual utility of using act $\act$ for state $\state$.  The expected utility
for the receiver of choosing act $\act \in \Acts$, given a message $\messg \in \Messgs$ and a
sender strategy $\Sstrat$ is the sum, for each state $\state \in \States$, of the prior
probability of state $\state$ occurring, times the probability the sender will choose message
$\messg$ given state $\state$, times the actual utility of using act $\act$ for state $\state$.

The \emph{fitness} of a behavioral strategy at a choice point is the
frequency-weighted average of expected utilities of each choice, given
the opponent's strategy:
\begin{align*}
  \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
\EU(\messg, \state,\Rstrat) \\
\Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
\cdot \EU(\act, \messg,\Sstrat)\,.
\end{align*}
The discrete-time replicator dynamic maps current strategies $\Sstrat$
and $\Rstrat$ to future strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$
in such a way that changes in frequency are proportional to expected
utilities. For behavioral strategies, the changes take place locally
at each of the agents' choice points:
\begin{align*}
  \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
    \EU(\messg, \state,\Rstrat)} {\Phi(\state,\Sstrat, \Rstrat)} \\
    \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
    \EU(\act, \messg,\Sstrat)} {\Phi(\messg,\Sstrat, \Rstrat)}  \,.
\end{align*}
The \rd in behavioral strategies describes the most likely path of development of average
behavior in a population of agents when behavioral adaptations occur independently at each
choice point, e.g., when agents adopt the behavior of another agent \emph{at a given choice
  point} if that ``local'' behavior is ``locally'' better than their own (see
\cite{Sandholm2010:Population-Game} for agent-level interpretations of the replicator dynamic
and \citeauthor{Cressman2003:Evolutionary-Dy} \citeyear{Cressman2003:Evolutionary-Dy} for
applications to extensive form games and behavioral strategies).

@@

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_ind3098.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0$}
    \label{fig:example_stratsA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_ind3452.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0.1$}
    \label{fig:example_stratsB}
  \end{subfigure}

@@

Representative examples for resulting strategy pairs are given in
Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA} shows a strategy pair at
stopping time with 90 states, tolerance $\toler = 0.1$ and imprecision $\imprecision = 0$. Zero
imprecision means that the trial was effectively an application of the plain \rd. Noteworthily,
the given sender strategy approximates a pure sender strategy that crisply partitions the state
space into non-convex sets. The irregular shape of the receiver strategy suggests that the
pictured strategy pair has not yet reached a dynamically stable state under the \rd. Indeed,
the trial was stopped for reaching the maximum of 200 rounds. In contrast, the outcome of a
trial with identical parameters, except with imprecision $\imprecision = 0.1$, which is shown
in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99 rounds of
iterated \rdd. The sender strategy shows a smooth blending from one ``category'' to the other,
and the receiver's interpretations are rather extended curves, peaking at a central point in
the relevant ``categories.''

@@

\paragraph{Entropy.} This classic information-theoretic notion
captures the amount of uncertainty in a probability
distribution. Roughly put, entropy of a signaling strategy captures
inverse distance from a pure strategy. The usual definition of entropy
applies directly to mixed strategies, but provably equivalent metrics
for behavioral strategies are ready to hand:
\begin{align*}
  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
  \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
  \state)) \\
  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
  \Rstrat(\state \probbar \messg) \cdot \log(\Rstrat(\state \probbar
  \messg)) \,. 
\end{align*}
Values obtained by these definitions are lower bounded by $0$ and
upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
\cdot \log(|\Messgs|)$ and $\log(|\States^\Messgs|) = |\Messgs| \cdot
\log(|\States|)$. We work with normalized values. The sender
strategies in Figures~\ref{fig:example_stratsA} and
\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.21$,
respectively. The receiver strategies have respective entropies $0.43$
and $0.84$. In general, we expected that vague languages would have
higher entropy than crisp ones and that increasing imprecision would
lead to increased entropy, all else being equal.

@@

\paragraph{Expected utility.} We also recorded the expected utility of
a strategy pair:
\begin{align*}
  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
\end{align*}
To make direct comparisons across different parameter settings, we
normalize expected utility by the maximal amount of expected utility
obtainable in the relevant game. The strategy pair in
Figure~\ref{fig:example_stratsA} has a normalized expected utility of
$0.992$, the pair in Figure~\ref{fig:example_stratsB} has
$0.901$. Generally, vagueness and imprecision can be expected to
decrease expected utility
\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is
whether communicative success drops unacceptably fast with moderate
levels of vagueness and imprecision.

>>>>

\newcommand{\rd}{\acro{rd}} % replicator dynamic
\newcommand{\rmd}{\acro{rmd}} % replicator mutator dynamic
\newcommand{\rdd}{\acro{rdd}} % replicator diffusion dynamic
\newcommand{\RD}{\ensuremath{\mathrm{RD}}} % replicator dynamic
\newcommand{\RDD}{\ensuremath{\mathrm{RDD}}} % replicator diffusion dynamic
\newcommand{\RMD}{\ensuremath{\mathrm{RMD}}} % replicator mutator
\newcommand{\icid}{\acro{icid}} % imprecise conditional imitation dynamic
\newcommand{\ICID}{\ensuremath{\mathrm{ICID}}} % imprecise conditional imitation dynamic

@@

\newcommand{\myred}[1]{\textcolor{red}{#1}}

@@

\begin{abstract}
  Signaling games are popular models for studying the evolution of meaning, but typical
  approaches do not incorporate vagueness as a feature of successful signaling.  Complementing
  recent like-minded models, we describe an aggregate population-level dynamic that describes a
  process of imitation of successful signaling behavior under imprecise perception and
  realization of similar stimuli. Applying this new dynamic to a generalization of Lewis'
  signaling games, we show that stochastic imprecision leads to vague, yet by-and-large
  efficient signal use, and, moreover, that it unifies evolutionary outcomes and helps avoid
  suboptimal categorization. \myred{The upshot of this is that generalization emerges, at an
    aggregate level, without agents actually generalizing.}
\end{abstract}

@@

A number of authors have recently tried to explain why vagueness evolved as something that is
itself useful
\citep[e.g.][]{Jaegherde-Jaegher2003:A-Game-Theoreti,Deemter2009:Utility-and-Lan,BlumeBoard2013:Intentional-Vag}.
Others have argued that vagueness is a natural byproduct of limitations in information
processing \citep[e.g.][]{FrankeJager2010:Vagueness-Signa} or the result of particular
low-level learning strategies \citep[e.g.][]{OConnor2013:The-Evolution-o}. This paper
contributes to the latter line of thought. Concretely, we introduce an dynamic that describes
average behavior resulting from imitation under confusability of similar stimuli. The dynamic
proposed here complements previous like-minded accounts. We show that stochastic noise can not
only lead to vague, yet relatively efficient signal use, but can also unify evolutionary
outcomes and help avoid suboptimal categorization. Moreover, the system shows generalization
behavior without a single generalizing agent.

The next section introduces the background against which the work presented here can be
appreciated. Section~\ref{sec:repl-diff-dynam} introduces our noise-affected imitation dynamic,
which is then explored in Section~\ref{sec:exploring-rdd}.  Section~\ref{sec:discussion}
reflects and compares our approach to others. % Two appendices cover technical material that is
% not strictly necessary to appreciate the main conceptual points of the paper.
% Appendix~\ref{sec:diffusion-as-special} shows that our replicator diffusion dynamic is a
% special case of the well-known replicator mutator dynamic. Appendix~\ref{sec:deriv-rdd-imit}
% gives a derivation of our dynamic as the most-likely path of a population whose agents imitate
% successful behavior without being aware of perceptual noise themselves.

@@

\section{Imprecise conditional imitation}
\label{sec:repl-diff-dynam}

Signaling agents can adapt their dispositions to act, given some feedback about their past
success, in multiple ways. Usually, we would assume that changes in behavior should, at least
on average, lean towards increasing chances of communicative success. Such a population
of signaling agents and their behavioral adaptations can be described at different levels of
abstraction. At the level of individual agents, we can describe a more or less idealized
process of how each agent adapts dispositions for future actions based on detailed descriptions
of information available to the agent, e.g., about the game, the other players' behavior and
the like. More abstractly, at the level of a population of agents, we can describe how
average behavioral dispositions will evolve. The latter population-level perspective abstracts
over petty stochastic fluctuations and zooms in on the general tendency or direction of
evolution that ensues from some (often: simplified) behavior at the agent-level.

We are interested here in exploring how a tendency to favor communicatively successful
strategies is affected by imprecise perception and realization of states. We therefore look at
a population-level dynamic that describes the most likely evolutionary path of a population of
signaling agents who imitate other agents' behavior (proportional to how successful it is) but
may occasionally confuse states for one another (proportional to how similar they are). We call
our dynamic \emph{imprecise conditional imitation dynamic} (\icid), because it derives from an
update protocol based on conditional imitation and allows for imprecise perception and
realization of states. In the special limiting case where imprecision vanishes, the process is
just the well-known replicator dynamic. To appreciate the \icid and its motivation, we first
briefly review the replicator dynamic and its derivation from conditional imitation of
successes.

@@

Fix a signaling game with finite states $\States$, messages $\Messgs$ and acts $\Acts$. Let
$\Pr(\cdot) \in \Delta(\States)$ be the prior distribution over states and
$\Util_{\sen,\rec} \mycolon \States \times \Messgs \times \Acts \rightarrow \mathds{R}$ the
sender's and receiver's utility functions. A \emph{behavioral strategy} is a function that maps
an agent's choice points to a probability distribution over available choices.\footnote{Our
  focus is on behavioral strategies, not \emph{mixed strategies}, i.e., probability
  distributions over functions from each choice point to an act, such as
  $s \in \Delta(\Messgs^{\States})$.  Dynamics on behavioral strategies assume that agents can
  adapt their behavior locally, so to speak, i.e., independently at each choice point. Our
  focus on behavioral strategies greatly reduces the complexity of the dynamic and simplifies
  numerical simulations. But it also seems the more plausible choice for imitation-based update
  protocols: agents only observe how, on some occasion, some other agent behaved in \emph{one}
  particular situation, not how that agent would behave in \emph{all} relevant choice
  situations; they imitate the use of a single word, so to speak, not a whole lexicon. (See
  \citet{Cressman2003:Evolutionary-Dy} for more on the difference between dynamics on mixed or
  behavioral strategies.)} The sender's behavioral strategies are functions
$\Sstrat \in \Delta(\Messgs)^\States$, thus mapping each state $\state \in \States$ to a
probability of each message $\messg \in \Messgs$ being sent; the receiver's are functions
$\Rstrat \in \Delta(\Acts)^\Messgs$, thus mapping each message $\messg \in \Messgs$ to a
probability of each state $\state \in \States$ being interpreted by the receiver given
$\messg$. Although behavioral strategies are probabilistic, we may conveniently imagine that
every individual agent has a non-probabilistic strategy nonetheless. If the population is large
enough, the number $\Sstrat(\messg \probbar \state)$, for instance, is then the probability
that a randomly sampled sender would have a strategy that responds with $\messg$ to $\state$.

The \emph{expected utility} of choices at each choice point is:
\begin{align*}
  \EU(\messg, \state, \Rstrat) & = \sum_{\act \in \Acts}
  \Rstrat(\act \probbar \messg) \cdot U_\sen(\state, \messg, \act) \\
  \EU(\act, \messg, \Sstrat) & = \sum_{\state \in
    \States} \Pr(\state) \cdot \Sstrat(\messg \probbar \state) \cdot
  U_\rec(\state, \messg, \act) \,.
\end{align*}
This means that the expected utility for the sender of choosing message $\messg \in \Messgs$,
given tate $\state \in \States$ and a receiver strategy $\Rstrat$ is the sum, for each act
$\act \in \Acts$, of the probability that the receiver will choose act $\act$ given message
$\messg$ times the actual utility of using act $\act$ for state $\state$.  The expected utility
for the receiver of choosing act $\act \in \Acts$, given a message $\messg \in \Messgs$ and a
sender strategy $\Sstrat$ is the sum, for each state $\state \in \States$, of the prior
probability of state $\state$ occurring, times the probability the sender will choose message
$\messg$ given state $\state$, times the actual utility of using act $\act$ for state $\state$.

% The \emph{fitness} of a behavioral strategy at a choice point is the
% frequency-weighted average of expected utilities of each choice, given
% the opponent's strategy:
% \begin{align*}
%   \Phi(\state,\Sstrat, \Rstrat) & = \sum_{\messg} \Sstrat(\messg \probbar \state) \cdot
% \EU(\messg, \state,\Rstrat) \\
% \Phi(\messg,\Sstrat, \Rstrat) & = \sum_{\act} \Rstrat(\act \probbar \messg)
% \cdot \EU(\act, \messg,\Sstrat)\,.
% \end{align*}

The discrete-time replicator dynamic maps current strategies $\Sstrat$ and $\Rstrat$ to future
strategies $\RD(\Sstrat)$ and $\RD(\Rstrat)$ in such a way that changes in frequency are
proportional to expected utilities:
\begin{align}
  \label{eq:RDDiscrete}
  \RD(\Sstrat)(\messg \probbar \state) & = \frac{\Sstrat(\messg \probbar \state) \cdot
    \EU(\messg, \state,\Rstrat)} {\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot
   \EU(\messg', \state,\Rstrat)} \\
    \RD(\Rstrat)(\act \probbar \messg) & = \frac{\Rstrat(\act \probbar \messg) \cdot
    \EU(\act, \messg,\Sstrat)} {\sum_{\act'} \Rstrat(\act' \probbar \messg)
    \cdot \EU(\act', \messg,\Sstrat)}  \,. \nonumber
\end{align}
Intuitively put, this definition updates the average frequency with which each choice (message
or act) will occur, independently at each choice point (state or message). For a given choice
point, say a state $\state$, the probability of seeing $\messg$ played by an average agent in
the population \emph{before} the update is $\Sstrat{\messg \probbar \state}$. \emph{After} the
update it will be proportional to the frequency before the update times the expected utility of
$\messg$ at state $\state$. The denominator is then a normalizing constant to make sure that we
obtain probabilities that sum to one from this. In particular, the denominator in the
definition above captures the average expected utility of all choices at the relevant choice
point.

\subsection{Conditional imitation of success}
\label{sec:cond-imit-succ}

The replicator dynamic is an abstract population-level dynamic that describes the mean,
to-be-expected change in behavioral dispositions in a population of signalers. There are
several ways of deriving the replicator dynamic from agent-level processes of behavioral
adaptation. We focus here on one of the simplest: \emph{conditional imitation of success}
\citep[see][for this and many other ways of deriving the replicator
dynamics]{Sandholm2010:Population-Game}. We choose this agent-level derivation, because it is
more convenient than its alternatives when we want to add imprecision in the following. The
main idea is that agents imitate the behavior of another agent at some choice point with a
probability that is proportional to the expected utility of the latter agent's act for that
choice point. Since the sim-max games that we will be looking at have all expected utilities
conveniently lie between 0 and 1, we may even identify the switching probability with the
expected utility directly.

Let's consider sender strategies, as the receiver case is parallel (as long as there is no
noise or imprecision). Revision by imitation of success proceeds as follows. Call
(misleadingly!) the agent who gets a chance to change behavior ``learner'' and the possibly
to-be-imitated agent ``teacher.'' A random learner is drawn from the population and given a
chance to change behavior at choice point $\state$. The probability that that learner plays
$\messg$ is $\Sstrat(\messg \probbar \state)$. The learner observes what a randomly sampled
teacher does at $\state$. That would be $\messg'$ with probability $\Sstrat(\messg' \probbar
\state)$. The learner then starts using $\messg'$ instead of $\messg$ with probability
$\EU(\messg', \state, \Rstrat)$. (Of course, $\messg'$ and $\messg$ could be the same.)

If agents get repeated updating chances for their choice points, the \emph{expected change} of
frequency for $\messg$ at $\state$ becomes:
\begin{align}
  \label{eq:MeanChange}
  \dot{\Sstrat}(\messg \probbar \state) = P(\messg' \rightarrow m, \state) - P(\messg \rightarrow m', \state)\,,
\end{align}
where $P(\messg' \rightarrow m, \state)$ is the ``inflow'' probability that agents switch from
$\messg'$ to $\messg$ and $P(\messg \rightarrow m', \state)$ is the ``outflow'' probability
that agents switch from $\messg$ to $\messg'$. Since we are dealing with expectations in a huge
population, these probabilities can be given as:
\begin{align*}
  P(\messg' \rightarrow m, \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg' \probbar
    \state)}_{\text{learner plays $\messg'$}} \cdot
  \underbrace{\Sstrat(\messg \probbar \state)}_{\text{teacher plays $\messg$}} \cdot
  \underbrace{\EU(\messg, \state, \Rstrat)}_{\text{EU teacher choice}} \\
  P(\messg \rightarrow m', \state) & = \sum_{\messg'} \underbrace{\Sstrat(\messg \probbar
    \state)}_{\text{learner plays $\messg$}} \cdot
  \underbrace{\Sstrat(\messg' \probbar \state)}_{\text{teacher plays $\messg'$}} \cdot
  \underbrace{\EU(\messg', \state, \Rstrat)}_{\text{EU teacher choice}}
\end{align*}

From this, we can simplify the expression of expected change in Equation~(\ref{eq:MeanChange})
to:
\begin{align*}
  \dot{\Sstrat}(\messg \probbar \state) & = \Sstrat(\messg \probbar \state) \cdot \EU(\messg,
  \state, \Rstrat) - \Sstrat(\messg \probbar \state) \cdot  \sum_{\messg'}
      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat) \\
& = \underbrace{\Sstrat(\messg \probbar \state)}_{\text{frequency of $\messg$ at $\state$}}  ( \underbrace{\EU(\messg,
  \state, \Rstrat)}_{\text{EU of $\messg$ at $\state$}} -   \underbrace{\sum_{\messg'}
      \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',\state,\Rstrat)}_{\text{average
      EU at choice point $\state$}}  ) 
\end{align*}
This is the continuous-time formulation of the replicator dynamics for behavioral
strategies. The discrete time formulation in Equation (\ref{eq:RDDiscrete}) solves this
continuous-time formulation by assuming that discrete update steps are infinitesimally
small:\footnote{Notice that we drop the denominator here, to obtain the non-payoff corrected
  replicator dynamic.}
\begin{align*}
  \dot{\Sstrat}(\messg \probbar \state) & = \RD(\Sstrat)(\messg \probbar \state) -
  \Sstrat(\messg \probbar \state) \\
  & = \frac{\Sstrat(\messg \probbar \state) \cdot \EU(\messg, \state,\Rstrat)} {\sum_{\messg'}
    \Sstrat(\messg' \probbar \state) \cdot
    \EU(\messg', \state,\Rstrat)} - \Sstrat(\messg \probbar \state) \\
  & = \frac{\Sstrat(\messg \probbar \state) \cdot \EU(\messg, \state,\Rstrat) - \Sstrat(\messg
    \probbar \state) \cdot \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
    \state,\Rstrat)} {\sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot
    \EU(\messg', \state,\Rstrat)}  \\
  & \propto \Sstrat(\messg \probbar \state) \cdot \left ( \EU(\messg, \state,\Rstrat) - \sum_{\messg'} \Sstrat(\messg' \probbar \state) \cdot \EU(\messg',
  \state,\Rstrat) \right )
\end{align*}

\bigskip

 Suppose, she currently plays $c$ at $k$. She
observes what a randomly sampled teacher would do at choice point $k$, say $c'$ and switches
from $c$ to $c'$ with switch probability $\text{SWITCH}(c \rightarrow c' , k)$, which we can
here identify with the expected utility $\EU(c',k)$ of $c'$ at $k$. The mean change induced by
offering all agents imitation chances for $k$ is:
\begin{align*}
  \text{mean change of frequency $c$, at $k$} & = \text{probability of switchers $c'
    \rightarrow c$ at $k$ \mygray{(inflow)}} \\
& - \text{probability of switchers $c \rightarrow c'$ at $k$ \mygray{(outflow)}}
\end{align*}
We should spell out the inflow and outflow here with some care, so as to prepare ourselves for
the subsequent addition of imprecise perception and realization of stimuli. So in general, we
would define:
\begin{align*}
  \text{inflow} & = \sum_{c'} \text{prob.~random learner maps $k$ to $c'$} \\
  & \ \ \ \times \text{prob.~learner observes $c$ by random teacher} \\
  & \ \ \ \times \EU(c', k) \\
  \text{outflow} & = \sum_{c'} \text{prob.~random learner maps $k$ to $c$} \\
  & \ \ \ \times \text{prob.~learner observes random teacher play $c'$ at $k$} \\
  & \ \ \ \times \text{SWITCH}(c \rightarrow c' , k)
\end{align*}

Suffice it to look at the case of a sender.

@@

  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_ind3098.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0$}
    \label{fig:example_stratsA}
  \end{subfigure}
  \hfill
  \begin{subfigure}[]{0.45\textwidth}
    \includegraphics[width=\textwidth]{plots/strat_example_ind1.pdf}
    \caption{$\ns = 90$, $\toler = 0.1$, $\imprecision = 0.0.5$}
    \label{fig:example_stratsB}
  \end{subfigure}

@@

Representative examples for resulting strategy pairs are given in
Figure~\ref{fig:example_strats}. Figure~\ref{fig:example_stratsA} shows a strategy pair at
stopping time with 90 states, tolerance $\toler = 0.1$ and imprecision $\imprecision = 0$. Zero
imprecision means that the trial was effectively an application of the plain \rd. Noteworthily,
the given sender strategy approximates a pure sender strategy that crisply partitions the state
space into non-convex sets. The irregular shape of the receiver strategy suggests that the
pictured strategy pair has not yet reached a dynamically stable state under the \rd. Indeed,
the trial was stopped for reaching the maximum of 200 rounds. In contrast, the outcome of a
trial with identical parameters, except with imprecision $\imprecision = 0.0.5$, which is shown
in Figure~\ref{fig:example_stratsB}, had converged (in our technical sense) after 99 rounds of
iterated \rdd. The sender strategy shows a smooth blending from one ``category'' to the other,
and the receiver's interpretations are rather extended curves, peaking at a central point in
the relevant ``categories.''

@@

\paragraph{Entropy.} This classic information-theoretic notion
captures the amount of uncertainty in a probability
distribution. Roughly put, entropy of a signaling strategy captures
inverse distance from a pure strategy. The usual definition of entropy
applies directly to mixed strategies, but provably equivalent metrics
for behavioral strategies are ready to hand:
\begin{align*}
  E(\Sstrat) = -\sum_{\state \in \States} \sum_{\messg \in \Messgs}
  \Sstrat(\messg \probbar \state) \cdot \log(\Sstrat(\messg \probbar
  \state)) \\
  E(\Rstrat) = -\sum_{\messg \in \Messgs} \sum_{\state \in \States}
  \Rstrat(\state \probbar \messg) \cdot \log(\Rstrat(\state \probbar
  \messg)) \,. 
\end{align*}
Values obtained by these definitions are lower bounded by $0$ and
upper bounded by, respectively, $\log(|\Messgs^\States|) = |\States|
\cdot \log(|\Messgs|)$ and $\log(|\States^\Messgs|) = |\Messgs| \cdot
\log(|\States|)$. We work with normalized values. The sender
strategies in Figures~\ref{fig:example_stratsA} and
\ref{fig:example_stratsB} have entropy $1.19e^{-5}$ and $0.08$,
respectively. The receiver strategies have respective entropies $0.43$
and $0.81$. In general, we expected that vague languages would have
higher entropy than crisp ones and that increasing imprecision would
lead to increased entropy, all else being equal.

@@

\paragraph{Expected utility.} We also recorded the expected utility of
a strategy pair:
\begin{align*}
  \EU(\Sstrat,\Rstrat \myts ; \myts \toler) = \sum_{\state \in
    \States} \sum_{\messg \in \Messgs} \sum_{\state' \in \States}
  \Pr(\state) \cdot \Sstrat(\state, \messg) \cdot \Rstrat(\messg,
  \state') \cdot \Util(\state,\state' \myts ; \myts \toler)\,.
\end{align*}
To make direct comparisons across different parameter settings, we
normalize expected utility by the maximal amount of expected utility
obtainable in the relevant game. The strategy pair in
Figure~\ref{fig:example_stratsA} has a normalized expected utility of
$0.99$, the pair in Figure~\ref{fig:example_stratsB} has
$0.95$. Generally, vagueness and imprecision can be expected to
decrease expected utility
\citep[c.f.][]{Lipman2009:Why-is-Language}. The crucial question is
whether communicative success drops unacceptably fast with moderate
levels of vagueness and imprecision.
